---
title: "ID in basketball"
author: 
- "Edgar Santos Fernandez"
- "Francesco Denti"
- "Kerrie Mengersen"
- "Antonietta Mira"
date: "2/6/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This file contains intructions to reproduce some of the results in the article
"The role of intrinsic dimension in high-resolution player tracking data - Insights in basketball".



```{r, eval = T, echo =F}
.libPaths('C:\\1\\R')
```

```{r, eval = T, message=F, warning=F}
.libPaths('C:\\1\\R')
library(data.table)
library(Rcpp)
library(RcppArmadillo)
library(ggplot2)
library(mvtnorm)
library(dplyr)
library(reshape2)
library(plot3D)
library(Matrix)
library(superheat)
library(mcclust)
library(RCurl)
library(jsonlite)
library(sp)
library(zoo)
library(lubridate)
library(stringr)
library(gridExtra)
library(grid)
library(RColorBrewer)
library(ggrepel)
library(viridis)
```

Hidalgo is implemented in C++ and the sampler and other functions are located in the file ```Hidalgo_Code.cpp```. 
Other useful functions from files ```Helpers_Code.R```  and ```plot functions.R``` need also to be sourced. 


```{r, eval = T, message=F, warning=F}
Rcpp::sourceCpp("Hidalgo_Code.cpp")
source("Helpers_Code.R")
source('plot functions.R')
```

The most of the movement data from the game is on the file ```CLEatGSW.RDS```.

```{r, eval = T}
allm <- readRDS("CLEatGSW.RDS")
allm$descrip <-ifelse(!is.na(allm$HOMEDESCRIPTION), as.character(allm$HOMEDESCRIPTION), as.character(allm$VISITORDESCRIPTION))

# selecting shot made and missed
allm <- dplyr::filter(allm, EVENTMSGTYPE %in% c(1,2) )


select_id <- function(all.movements = all.movements, id = id){
  e1 <- dplyr::filter(all.movements, event.id == id) 
  e1$x_loc <- as.numeric(as.character(e1$x_loc))
  e1$y_loc <- as.numeric(as.character(e1$y_loc))
  e1$shot_clock <- as.numeric(as.character(e1$shot_clock))
  e1$shot_clock2 <- round(e1$shot_clock,1)
  e1 <- dplyr::arrange(e1, datetime2)
  
  e1$ii <- 1
  e1 <- e1 %>% dplyr::mutate(act = ifelse(shot_clock2 > lag(shot_clock2), ii + 1,ii)) # need to split actions because multiple actions are mingled
  
  breaks <- c(ymd_hms('2010-01-01 00:00:00'), e1[which(e1$act == 2),]$datetime2, Sys.time())
  e1$action <- cut.POSIXt(e1$datetime2, breaks = breaks, labels = 1:(length(breaks)-1))
  
  e1 <- dplyr::filter(e1, shot_clock2 < 24)
  e1 <- e1 %>% dplyr::group_by(event.id) %>%  dplyr::mutate(ids = dplyr::row_number())
  
}


# table with the action to be used.
# this was done manually because many events have issues with the clock
# 
table_fixing_clock <- read.csv('across plays fixing clock issues.csv')
table_fixing_clock$quarter <- factor(table_fixing_clock$quarter)
table_fixing_clock$EVENTMSGTYPE <- factor(table_fixing_clock$EVENTMSGTYPE)

allm$EVENTMSGTYPE <- droplevels(allm$EVENTMSGTYPE)
allm$ball_possession <- factor(allm$ball_possession)

allm <- allm %>% left_join(table_fixing_clock, by = c('event.id', 'quarter', 'EVENTMSGTYPE', 'ball_possession'))

# will exclude the plays with issues according to the visualization done with assess

allm <- dplyr::filter(allm, comment != 'e')

```

```{r, eval = T}
unique_events <- allm %>% dplyr::select(event.id, EVENTMSGTYPE, ball_possession, quarter, descrip, SCORE) %>% distinct()

ev <- unique(allm$event.id) # events


# computing ID for event 6 
i = 6

s15 <- select_id(all.movements = allm, id = ev[i]) 
s15 <- filter(s15, action == table_fixing_clock[table_fixing_clock$event.id == ev[i],]$action)

if(unique(s15$quarter) %in% c(1,2) ){
  if(unique(s15$ball_possession) == 'home') { 
         s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc < 47)$datetime2))
         }
if(unique(s15$ball_possession) == 'away') {
         s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc > 47)$datetime2))
       }
    }


#filter offensive court only
if(unique(s15$quarter) %in% c(3,4) ){
      if(unique(s15$ball_possession) == 'home') { 
        s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc > 47)$datetime2))
      }
if(unique(s15$ball_possession) == 'away') {
        s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc < 47)$datetime2))
      }
}


```


```{r, eval = T}

stamps <- unique(s15$datetime2)[seq(1, length(unique(s15$datetime2)), by = 10)] 

s15 <- filter(s15, datetime2 %in% stamps) %>% dplyr::arrange(event.id, datetime2, team, lastname)
  
s15$team <- as.character(s15$team)
  #levels(s15$team) <- c('away', 'home', 'ball')
s15$team <- ifelse(is.na(s15$team) , 'ball', s15$team)
####################

  s15$x <- s15$x_loc
  s15$y <- s15$y_loc
  s15$player <-  rep(c(paste0('a', 1:5), paste0('h', 1:5), 'ball') , times = round(nrow(s15)/11))[1:nrow(s15)]

  s15 <- s15 %>% arrange(player, datetime2)
  s15$stamp <- rep(1:length(stamps),times = 11)
  
  s15 <- s15 %>% dplyr::group_by(player, lastname) %>% 
    dplyr::mutate(dist = sqrt( (x-lag(x)) ^ 2 + (y - lag(y)) ^ 2) ) #%>% dplyr::group_by(player, shot_clock) 

  if(unique(s15$ball_possession) == 'home') { 
     players <-  c("ball", unique(s15[s15$team == 'home',]$lastname))}
  
  if(unique(s15$ball_possession) == 'away') { 
    players <-  c("ball", unique(s15[s15$team == 'away',]$lastname))}
  
s15$lastname <- factor(s15$lastname)

colp <- brewer.pal(12, "Paired")[c(6, 2,4,8, 10,12)]

# plotting the movement of the players on attack
# https://youtu.be/jb57MFQLoRo?t=17

p1 <- P_180 + 
    geom_point(data = dplyr::filter(s15,lastname %in% players ), aes(x_loc, -y_loc, group = lastname, col = lastname, shape = team), size =1 ) + 
    geom_path(data = dplyr::filter(s15,lastname %in% players), aes(x_loc, -y_loc, group = lastname, col = lastname), size = 0.3) +
geom_text(data = dplyr::filter(s15,lastname %in% players), aes(x_loc, -y_loc+1, label = stamp, col = lastname), size = 1.5)+
    scale_color_manual(values = colp) +
    theme_bw() + ggtitle(paste0('event_id = ', unique(s15$event.id))) 

```
  
```{r, eval = T, cache = T}
dist_stamp <- s15 %>% dplyr::group_by(datetime2) %>%
    dplyr::summarize(dist = sum(dist, na.rm = T))
  
s15_ <- data.table::dcast(setDT(s15), event.id + quarter + EVENTMSGTYPE+
                EVENTMSGACTIONTYPE + SCORE + SCOREMARGIN + HOMEDESCRIPTION +
                VISITORDESCRIPTION + datetime2 + shot_clock + game_clock ~ player, value.var=c("x", "y", "radius"))
  
s15_ <- s15_ %>% dplyr::select(-c("radius_a1","radius_a2","radius_a3", "radius_a4","radius_a5","radius_h1",        
                                "radius_h2","radius_h3","radius_h4","radius_h5") )
  s15_ <- as.data.frame(s15_)
  
  vars <- c(names(s15_)[grepl('_h', names(s15_))] , names(s15_)[grepl('_a', names(s15_))] )  
  
  DATA <- s15_[, vars] 
```

Applying HIDALGO to the movement data with dimension D = 20. 
```{r, eval = T, cache = T}
  nsim = 20000 # simulations
  burn_in = 10000 # burn in period
  L      <- 3 
  Lista <- data.preprocessing(DATA, q = 3) # computes the mus
  out3R   <- HIDALGO(mu_obser = Lista[[1]],
                         L = L,
                         NSIM = nsim, 
                         burn_in = burn_in,
                         thinning = 1, 
                         verbose = F, verbose_step = 1000,
                         a0_d = 2, b0_d = 1, Nq = Lista[[2]],
                         q = Lista[[3]],CSI = .75, 
                         t1 = .5, t2 = .001,REPULSIVE = T,
                         alpha = rep(1,L),D = 20,TRUNC = F)
  tracked_d3R=Tracking(out3R$AC,out3R$AD, N = nrow(DATA), NSIM = nsim)

ID <- apply(tracked_d3R,1,median)
```

Plotting the ID accross time stamps:

```{r, eval = T, cache = T}    
df15 <- data.frame(id = ID, i = 1:length(ID))
 
p2 <- ggplot(df15) + 
  geom_line(aes(y = id, x = i, col = id), size = 1.5)+
  scale_color_viridis()+
  xlab('frame')+ ylab('ID')+
  theme_bw() 

```



```{r, eval = T}
grid.arrange(p1, p2, nrow = 2, ncol = 1, heights=c(10,5), widths=c(11))
```